Supplemental Material for ``Assessing United States county-level exposure for research on tropical cyclones and human health"

### 2020-09-09 ehp_revision/response_to_reviewers.Rmd

\input{tables/rain_misclass.tex}

\input{tables/flood_misclass.tex}

\input{tables/tornado_misclass.tex}

We have added four new tables in the Supplemental Material, one for each of the
hazard-based metrics we considered. We have reproduced these tables in this
response document as Tables \@ref(tab:misclasswind)--\@ref(tab:misclassflood)
Each breaks down agreement and disagreement between the distance-based proxy
metric and that hazard-based metric within several storms in our study period.
These tables focus on storms with at least 250 counties exposed by at least one
of the metrics considered. These tables allow readers both to investigate
general patterns in disagreement and exposure misclassification across these
storms and also to see cases of unusual storms. For example, for most storms,
disagreement between the distance proxy metric and the wind-based metric
resulted mainly from mis-classification of counties as exposed based on distance
while they did not experience gale-force winds. However, there were a few
storms, like Hurricane Bertha in 1996, where the inverse was the case, and a
distance-based metric would have missed a large number of counties that
experienced gale-force winds from the storm (misclassified them as unexposed),
but not misclassified many counties as exposed when they in fact did not
experience winds at this level.

### 2020-09-09 ehp_revision/response_to_reviewers.Rmd

\textit{Lines 185--187 from original manuscript:}

\begin{quotation}
\noindent
\textit{"We used tracking data from HURDAT2, which records the storm center’s
position at the synoptic times of 6:00 am, 12:00 pm, 6:00 pm, and 12:00 am
Coordinated Universal Time (UTC)."}
\end{quotation}
\end{shaded}

**Response:**

Some weather data are collected at standardized times, so that observations from
different locations can be meaningfully integrated to provide a view of
large-scale (synoptic) weather systems and patterns [@usnwstime;
@willoughby2007hurricane; @willis2006cleveland; @raines1996getting]. This
practice extends back over 100 years, when weather data for the US were
collected by local weather offices and sent by telegram at regular times each
day into a central office, which used this information to create synoptic
weather maps that captured major weather systems in the county and then
sent back forecasts to local weather offices, also at standardized times
[@willis2006cleveland; @raines1996getting]. Synoptic weather times are typically
set based on the Universal Time Constant, with times recorded in Coordinated
Universal Time (i.e., Greenwich Mean Time), and time stamps on weather maps and
data based on this time zone often is given a "Z" (for "Zulu", an earlier name
for this timezone) after the time [@usnwstime;
@morris2008time].

@@

\medskip

\begin{shaded}
\textbf{R3 Comment 6:}
Table 1: I suggest providing the rationale for picking the thresholds for
classifying continuous exposures into binary ones within the table. 
\end{shaded}

**Response:**

For all outcomes, there are various reasonable choices that can be picked for
the threshold in defining exposure. There is not room to go into these details
within Table 1, so instead we added a separate table to the Supplemental
Material that provides both some reasoning behind the thresholds selected in
this study and also other reasonable thresholds that could be considered.
This is Table S1 of the revised manuscript and has been reproduced as 
Table \@ref(tab:thresholds) at the end of this response document.

@@

\medskip

\begin{shaded}
\textbf{EHP Request 2:}
Main text figures (general)
To ensure that figures are accessible to readers with impaired color vision, do not use color as the sole means of conveying information in a figure unless absolutely necessary. Use different symbols, shading/textures, and line patterns instead of (or in addition to) color to distinguish among different data points. Use contrasting colors that can be easily distinguished from each other when the figure is printed in black and white. For details, see: \url{https://ehp.niehs.nih.gov/authors/figures}.
\end{shaded}

**Response:**

In all these figures, we have used colormaps that are both accessible to readers
with several forms of color vision deficiency, including the most common type
(deuteranomaly, a type of red-green color vision deficiency), and can be easily
distinguished when the figure is printed in black and white [@viridis;
@van2015mpl]. The colormaps we used are all multi-hue sequential maps that avoid
red-green contrasts, while maximizing dynamic range in comparison to something
single-hued, like grayscale [@nunez2018optimizing; @viridis]. They are all
perceptually uniform across the scale for those with normal color vision
[@liu2018somewhere] and are close to perceptually uniform for those with common
forms of color deficiency [@nunez2018optimizing]. They reproduce their original
sequential gradient when printed in grayscale and so are perceptable when
printed on a black-and-white printer [@van2015mpl; @nunez2018optimizing].
Previous research has found that colormaps in this family allow viewers to more
quickly and accurately judge relative distances in scientific figures
[@liu2018somewhere]. The family of colormaps we used in the figures is currently
considered the gold standard for scientific figures [@nunez2018optimizing].

To help illustrate, in Figures \@ref(fig:fig5check1) and \@ref(fig:fig5check2)
of this response document, we have reproduced the current version of Figure 5 as
it would appear to someone with three common types of color vision deficiencies:
deuteranomaly (defective green cone cells, the most common type), as well as
protanopia (defective red cone cells) and tritanopia (defective blue cone
cells). We have also shown how the Figure would appear in grayscale in Figure
\@ref(fig:fig5check2) (right). We have included similar versions of other
figures from the manuscript in response to later requests from EHP in this
response.

### 2020-09-09 ehp_revision/writing/ehp_response_darlings.tex
Work some more on response

R3 Comment 5

@@

>>>>

# R3 Comment 5

@@

# R3 Comment 6

- **Wind:** This threshold is the threshold for gale-force and higher winds
based on the Beaufort wind scale. This value is used by the National Hurricane
Center [?] as the outer limit of winds for its wind radii product in storm 
tracking data. Other reasonable choices for a threshold of wind exposure 
could include the two more intense thresholds also used in the NHC [?] wind
radii estimates of storm extent (50 knots and 64 knots, the later of which 
is the threshold for hurricane-force winds on the Beaufort Scale) or other
thresholds on the Beaufort Scale (e.g., "strong gale", which starts at 
41 knots; "storm", which starts at 48 knots; "violent storm", which starts
at 56 knots).

- **Rain:** This threshold is ... Other reasonable thresholds might be ... or
might incorporate different ranges of days around the storm's closest approach
to a county in calculating the cumulative precipitation. For example, ... For
the most part, using our window (two days before to one day after a storm's
closest approach) captured most of the storm-associated precipitation that would
have been counted using a wider window from five days before to two days after a
storm's closest approach, even for a large, slow storm. However, for a storm
that loops and so passes some counties twice, such as Allison (2001), this
four-day window may not be adequate, and lead to an underestimation of total
rainfall associated with the storm relative to estimates for faster-moving
storms. Other storms that looped over or near land and so passed close to US
counties twice include Allison (1989), Alberto (1994), Gordon (1994), Dennis
(1999), Edouard (2002), Fay (2002), Alex (2004), Ivan (2004), Dennis (2005), and
Ophelia (2005). One difficulty in measuring rain exposure consistently across
many different storms is how many days to include in the rain total. Some
storms, especially ones associated with dangerous rainfalls, can move very
slowly, meaning a window of several days needs to be used to fully capture
storm-related rainfall. For example, one study used a period of over four days
to provide storm rainfall totals from the slow-moving Hurricane Danny in 1997
[@medlin2007]. Single-storm studies often measure rainfall as the total rainfall
over the full "event" (e.g., @medlin2007); however, the timing of the event is
determine in a customized way for each storm, making it difficult to extend to
multi-storm analyses. We included more days prior to the storm because
storm-related rains often precede the passing of the storm's center [ref].

- **Distance:** This threshold is one that has been commonly used in studies in
which distance from the storm's track is used as a proxy for tropical cyclone
exposure (for example, [refs]). Further, [anything in relation to storm size?].
Other reasonable thresholds for distance include ..., all of which have also
been used in previous research in assessing tropical cyclone exposure using
distance from the storm's track as a proxy. Other distance-based exposure
metrics have defined "exposure" to occur only when a storm's center track passes
through a county's boundary (e.g., @zandbergen2009, who describe this metric of
exposure as a "hit" on the county; @kinney2008 is another example that
considered highest exposures for counties directly on the storm's path). This
way of measuring exposure can exclude nearby counties that suffered extreme
conditions from the storm but were not directly on the hurricane's track
[@kruk2010], especially since hurricanes can have a width ([typical storm width;
citation])[JMF-- Weatherford and Gray paper referenced in Grabich 2016 may be a
good place to start. Suggest 30-60km width of the eye of the storm itself] that
is much larger than a typical county's width and since a hurricane's maximum
winds can extend well away from the storm's center. Further, exposure assessed
using this method is associated with the size [@zandbergen2009; @kruk2010] and
shape [@zandbergen2009] of a county. (While some studies have avoided this
problem by determining exposure for equal-area grid cells (e.g., [@kruk2010]),
in studies where exposure needs to be matched with a county-level outcome, some
process of aggregation from grid cells to counties would still need to be
performed.) By contrast, assessing distance-based exposure based on measuring
distance between the storm track and each county's center would be less
susceptible to differences in county size and shape and so might be a more
reliable metric of exposure than determining if a storm track ever crossed a
county boundary. Further, assessing the minimum distance between county center
and the storm track, as we do here, allows for a continuous, rather than binary,
metric of distance-based exposure to a storm. To date, distance parameters
involved with assessing risk of a particular storm have been rather arbitrary,
contributing to the necessity in understanding how such parameters influence a
county's exposure status. In assessing the risk of a given storm based on
distance, recent studies have defined the distance from the storm track affected
by a given storm differently. @czajkowski2011 assessed county-level risk and
exposure based on a three-tiered definition, with primary counties being those
closest to the storm track on either side, secondary counties being adjacent to
primary counties, and tertiary counties adjacent to secondary counties. Such a
definition resulted in an exposure defenition based on an average distance
radius of 120 km [75 miles] on either side of the storm track [@czajkowski2011].
Such a distance is slightly larger than that commonly used by public health
departments (i.e., 100 km) [@czajkowski2011]. Other studies used distance
buffers of 30 kilometers [@currie2013] or of 30, 60, and 100 kilometers as
thresholds for identifying distance-based exposure [@grabich2015measuring] [check the
Grabich paper-- was this distance to any part of a county, or to the center of
the county?; JMF-- I think any part of the county]. Another study used a
distance criterion of 60 kilometers from the storm track, classifying all
counties for which the 60 km buffer fully or partially passed through the
county's boundaries as exposed [@grabich2015] [again, check to see if this was
to any point in the county or to some center point of the county][JMF-- it is
not abundantly clear in the Grabich 2016 article; based on their graphs it seems
like they considered a county exposed if the definition they used crossed any
county boundary. The Grabich 2015 paper specifies both full and partial counties
being included]. In a study assessing the association between hurricanes and
undesireable birth outcomes, researchers found that results were not sensitive
to the omission of residences 100 km from the storm path, and that results
varied insignificantly from 30-75 km [@currie2013]. One study used distance to
determine exposure to landfalling US hurricanes by assigning counties along the
storm's tracks as "primary" exposures, counties neighboring primary counties as
"secondary" counties, and counties neighboring secondary counties as "tertiary"
counties; this method correspond to a distance buffer of approximately 75 miles
(120 kilometers) from the storm track for a county to qualify as "exposed" to a
storm [@czajkowski2011]. The study found that about 30 direct hurricane
fatalities over their study period (1970--2007) occurred outside of this
distance range from the storm center [@czajkowski2011], suggesting that this
distance choice can cause some counties with dangerous exposures to a storm to
be mis-identified as "unexposed". Tropical cyclones vary dramatically in size:
US tropical cyclones have been observed with radii to maximum winds as small as
20-km and as large as 200-km [@mallin2006; @quiring2011variations].
